Efficient Constructions of Prefer-Max Sequences and of Infinite de Bruijn Sequences

### 2017-01-09 paper.tex
Fixed some of Gal's corrections

\documentclass{article}
%\documentclass[letterpaper, 10 pt, conference]{ieeeconf}

%\IEEEoverridecommandlockouts        % This command is only needed if you want to use the \thanks command
%\overrideIEEEmargins                        % See the \addtolength command later in the file to balance the column lengths

@@

%------------------------------------------- preamble ---------------------------

@@

\begin{algorithm}
Repeatedly append the maximal symbol over the alphabet $0,\dots,k-1$ such that the word formed by the $n$ most recent symbols is new. In the first $n$ steps, act as if thee are $n$ invisible zeros before the sequence. Stop after appending $n$ consecutive zeros. 
\caption{The $(k,n)$-prefer-max sequence.}
\label{pref-max}
\end{algorithm}

@@

An obvious limitation of the above construction is, of course, that, when $n$ and $k$ grow, the amount of time and memory needed to produce the next symbol grows exponentially. Therefore, this construction is not practical for generating de Bruijn sequences of large alphabets and spans. It is used here as a mathematical definition of the sequence. An efficient algorithm for constructing this sequence is one of the contributions of this paper.

@@

\begin{theorem}[Onion]
	The last $(k-1)^n+1$ symbols of the $(k,n)$-prefer-max sequence form the $(k-1,n)$-prefer-max sequence.
\label{thm:onion}
\end{theorem}

The observation made in the last theorem directed us to examine the reverse of the prefer-max sequence. Specifically, the structure identified in Theorem~\ref{thm:onion} means that, if we manage to construct the sequence backwards, we do not have to specify $k$ ahead of time. Instead, we can add symbols forever while $k$ grows in layers: we first write $0^n$, then a word whose substrings of length $n$ are $\{0,1\}^n$, then extend with other words in $\{0,1,2\}^n$, and so on.
1
The second contribution of this work is then an efficient construction of this reversed sequence. This is formulated in Algorithm~\ref{inf-construction}. We say that this construction is efficient because each step can be computed in $O(n^2)$ operations by enumerating all the $n$ rotations of $\sigma w$ and comparing them lexicographically (reading each word from right to left).

@@

While our main focus is on analyzing the reversed sequence, we also have a result about the prefer-max sequence itself. Our result gives the first (to the best of our knowledge) efficient construction of that sequence:

@@

Each step of Algorithm~\ref{efficient-pref-max} can be computed in $O(n^2)$ time because it suffices to examine for $\sigma'$ only symbols that are in $\sigma w$ or are smaller by one than a symbol in $\sigma w$, as we will establish later in the paper.
The fact that this is indeed the prefer-max sequence is stated in the following theorem:

@@

\section{A direct proof of the Onion theorem}

@@

\begin{repeatedtheorem}{thm:onion}[Onion]
	{The last $(k-1)^n$ symbols of the $(k,n)$-prefer-max sequence form the $(k-1,n)$-prefer-max sequence.}
\end{repeatedtheorem}
\begin{proof}
	Denote the symbols of the $(n,k$)-prefer-max sequence by
	$s_1,s_2,\dots,s_{k^n+1}$ and let, for notational convenience, $s_i=0$ for all
	$i<1$.  For each $i\geq 1$ let $w_i=s_{i-n+1}\cdots s_i$ be the ``sliding window''
	observed at the $i$th symbol of the sequence. 
	The proof goes by showing that $\{w_1,\dots,w_{k^n-(k-1)^n}\} = \{0,\dots,k-1\}^n \setminus \{0,\dots,k-2\}^n$.

@@

	The sequence is therefore as follows:
	$$\underbrace{s_1,\dots,s_{i_0-n-1},k-1,\overbrace{0,\dots,0}^\text{$n\text{-}1$ times}}_{\text{all the $w_i$s here contain
				$k-1$}},\underbrace{k-2}_{s_{i_0}},s_{i_0+1},\dots,s_{k^n+1}$$
	To complete the argument, we need to show that all substrings of length $n$ that
	contain $k-1$ appear before $s_{i_0}$.

>>>>

\documentclass{article} %\documentclass[letterpaper, 10 pt,
% conference]{ieeeconf}

%\IEEEoverridecommandlockouts        % This command is only needed if you want
%to use the \thanks command
%\overrideIEEEmargins                        % See the \addtolength command
%later in the file to balance the column lengths

@@

%------------------------------------------- preamble
%---------------------------

@@

	\begin{algorithm}
		Repeatedly append the maximal symbol over the alphabet
		$0,\dots,k-1$ such that the word formed by the $n$ most recent symbols is new.
		In the first $n$ steps, act as if there are $n$ invisible zeros before the
		sequence. Stop after appending $n$ consecutive zeros. 
		\caption{The $(k,n)$-prefer-max sequence.} 
		\label{pref-max} 
	\end{algorithm}

@@

	An obvious limitation of the above construction is, of course, that, when $n$
	and $k$ grow, the amount of time and memory needed to produce the next symbol
	grows exponentially. Therefore, this construction is not practical for
	generating de Bruijn sequences of large alphabets and spans. It is used here as
	a mathematical definition of the sequence. More formally, a shift rule for a de
	Bruijn sequence of order $n$ is a function that maps each word of length $n$ to
	the symbol that follows that word as a substring in sequence. The above is an
	inefficient shift rule for the prefer-max sequence. The first (to the best of
	our knowledge) efficiently computable shift rule for the prefer-max sequence is
	one of the contributions of this paper.

@@

	\begin{theorem}[Onion] 
		The $(k-1,n)$-prefer-max sequence is a suffix of the
		$(k,n)$-prefer-max sequence. 
		\label{thm:onion}
	\end{theorem}

	The observation made in the last theorem directed us to examine the reverse of
	the prefer-max sequence. Specifically, the structure identified in
	Theorem~\ref{thm:onion} means that, if we manage to construct the sequence
	backwards, we do not have to specify $k$ ahead of time. Instead, we can add
	symbols forever while $k$ grows in layers: we first write $0^n$, then a word
	whose substrings of length $n$ are $\{0,1\}^n$, then extend with other words in
	$\{0,1,2\}^n$, and so on. The second contribution of this work is then an
	efficient construction of this reversed sequence. This is formulated in
	Algorithm~\ref{inf-construction}. We say that this construction is efficient
	because each step can be computed in $O(n^2)$ operations by enumerating all the
	$n$ rotations of $\sigma w$ and comparing them lexicographically (reading each
	word from right to left).

@@

	While our main focus is on analyzing the reversed sequence, we also have a
	result about the prefer-max sequence itself. Our result gives the first (to the
	best of our knowledge) efficiently computable shift rule for that sequence:

@@

	Each step of Algorithm~\ref{efficient-pref-max} can be computed in $O(n^2)$
	time because it suffices to examine, as candidates for $\sigma'$, only symbols
	that are in $\sigma w$ or are smaller by one than a symbol in $\sigma w$, as we
	will establish later in the paper. The fact that this is indeed the prefer-max
	sequence is stated in the following theorem:

@@

	\section{Proofs}

@@

	\begin{repeatedtheorem}{thm:onion}[Onion] 
		The $(k-1,n)$-prefer-max sequence is a suffix of the $(k,n)$-prefer-max sequence. 
	\end{repeatedtheorem}
	\begin{proof} Denote the symbols of the $(n,k$)-prefer-max sequence by
		$s_1,s_2,\dots,s_{k^n+1}$ and let, for notational convenience, $s_i=0$ for all
		$i<1$.  For each $i\geq 1$ let $w_i=s_{i-n+1}\cdots s_i$ be the ``sliding
		window'' observed at the $i$th symbol of the sequence. The proof goes by
		showing that $\{w_1,\dots,w_{k^n-(k-1)^n}\} = \{0,\dots,k-1\}^n \setminus
		\{0,\dots,k-2\}^n$.

@@

		The sequence is therefore as follows:

		$$\underbrace{s_1,\dots,s_{i_0-n-1},k-1,\overbrace{0,\dots,0}^\text{$n\text{-}1$ times}}_{\text{all the $w_i$s here contain $k-1$}},\underbrace{k-2}_{s_{i_0}},s_{i_0+1},\dots,s_{k^n+1}$$ To complete the argument, we need to show that all substrings of length $n$ that contain $k-1$ appear before $s_{i_0}$.
